---
title: "Data Analysis"
author: "Vikram Vijayabaskar"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 2
    embed-resources: true
pagetitle: "UCR STAT107 Final Data Analysis Project - {{< meta title >}}"
editor: visual
---

# Introduction {.unnumbered}

## Introducing the data

The data set I will be using is the Pokemon Go evolution data set. It includes 75 Pokemon evolutions and around 27 variables, as well as information about each Pokemon before and after evolution (combat power, hit points, size, power up costs, etc). Evolution in Pokemon means a Pokemon is growing stronger, thus, in this project, I'll be analyzing/predicting post evolution combat power.

## Code Book

(All descriptions are from the readme attached to the dataset)

| Variable | Description | Why it is used |
|------------------------|------------------------|------------------------|
| `species` | The Pokémon's type, e.g. Pidgey. | I use this variable to compare whether CP (combat power) differs between species |
| `cp` | Pre-evolution Combat Power, which is a summary of the Pokémon's strength for battling prior to the evolution of the Pokémon. | This is the main variable I will be using to predict. |
| `cp_new` | Post-evolution Combat Power. | This is the main variable I will be PREDICTING. |
| `hp` | Pre-evolution Hit Points, which is a summary of how difficult it is to weaken the Pokémon in a battle. | I use this variable for background summary statistics |
| `hp_new` | Post-evolution Hit Points. | I use this for comparing background statistics after evolution. |

## Omitted Variables

I will not be omitting any variables from the data set, HOWEVER, my main model will only be focusing on the variables that directly answer my research question. The model focuses on `cp_new`, `cp`, and the other variables left in the code book. Other variables in the data set such as `attack`, `attack types`, `weight`, `height`, and more aren't included in the main regression model because they are not integral to our research question and might skew the data.

## Motivation

As a kid growing up with Pokemon because of an older brother, who was much more fascinated by the stats of Pokemon because he wanted the strongest team possible, I was interested in being able to predict whether or not I can tell a Pokemon is going to have a strong evolution simply based off its pre-evolution, therefore (even though I personally don't play Pokemon anymore) I could decide which Pokemon I would want based off their pre-evolution and end up with the strongest team I could have. Some background information that suggests that there's reason to investigate this is because sometimes some Pokemon have very weak pre-evolutions that evolve into very strong Pokemon, I want to be able to predict that.

## Research Question

My research question is: Can we predict a Pokemon's post evolution combat power using it's pre-evolution combat power, and does this relationship differ by species?

## Hypothesis

My hypotheses are

$H_0$ - Pre-evolution Combat Power is not associated with Post-Evolution Combat Power.

$H_A$ - Higher Pre-Evolution Combat Power is associated with higher Post-Evolution Combat Power.

# Methods

## Preliminary Summaries

Before diving into models, I will summarize the data by finding the total sample size, the number of variables, and summary statistics for the main outcome, which is the variable `cp_new`. I will also find the number of Pokemon in each `species` group. This shows how much ndata is available for each each species before comparing models

## Main Model

The main method I will be using for this project/analysis is linear regression.

$cp\_new = \beta_0 + \beta_1cp + \beta_2species + \epsilon$

This model lets me test whether pre-evolution CP is associated with post-evolution CP while also accounting for differences between Pokemon species.

## Main Figure

The main figure of this analysis will be a scatterplot of `cp_new` versus `cp`, colored by `species`, with a fitted regression line. This figure shows us the relationship between the two variables we are comparing, pre-evolution CP and post-evolution CP.

## Model assumptions

The assumption the model takes is that the relationship between `cp` and `cp_new` is approximately linear, observations are independent, residual variance is roughly constant, and that the residuals are normal. I will check these assumptions during my analysis but the report will focus on the main figure.

##Software and functions I will be using
I'll be using `tidyverse` for data cleaning, summary tables, and graphics. I will also use `lm()` to fit the regression model, `summary()` to view model results, and `ggplot2` to create the figure.

# Results

```{r}
#loading packages
library(tidyverse)
```
```{r}
#load data
pokemon_go <- read_csv("pokemon_go.csv")

#clean variables
pokemon_go <- pokemon_go |>
  mutate(species = factor(species))
```
# Data Summary
```{r}
pokemon_go |>
  summarize(n_rows = n(),
            n_variables = ncol(pokemon_go),
            mean_cp_new = mean(cp_new),
            sd_cp_new = sd(cp_new),
            median_cp_new = median(cp_new),
            iqr_cp_new = IQR(cp_new))
```
```{r}
#counts the amount of each species
pokemon_go |>
  count(species)
```
## Fitting

```{r}
main_model <- lm(cp_new ~ cp + species, data = pokemon_go)

summary(main_model)
```

